Evaluating sensitivity to measurement error in dissolved gas measurements

### 2024-04-22 scripts/DG_sensitivity_to_measurement_error.Rmd
Added more precise methods to simulations

<style>
.vscroll-plot {
    width: 800px;
    height: 600px;
    overflow-y: scroll;
    overflow-x: hidden;
}
</style>

# Background
In this document we explore a range of models for fitting dissolved $\text{N}_2\text{O}$ data observed as part of the 2017 National Lakes Assessment (NLA).

# Setup R
```{r setup_1, include=FALSE, echo=FALSE}
library(knitr)
# opts_knit$set(root.dir = "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/scripts")
# 
# #set directories and R package library
# setwd("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/scripts")
# .libPaths("C:/Users/rmartin/Desktop/R/library")
```

@@

# Set up model for simulating true value

@@

B <- 99 # barometric pressure (kPa)
V_g <- 35 # volume of reference gas in headspace
V_w <- 105 # volume of water below headspace
T_c <- 23 # temperature ( C )
C_e <- 0.307 # gas mixture ratio in headspace
C_a <- 0.310 # gas mixture ratio in reference
H <- 0.00024 #mol m-3 Pa, range: 0.00018 - 0.00025; Henry's law constant

C <- 1e-6 * B * ( V_g * ( C_e - C_a ) / ( 8.3144598 * ( T_c + 273.15 ) * V_w ) + H * exp( 2700 * ( 1 / ( T_c + 273.15 ) - 1 / 298.15 ) ) * C_e )

@@

```
## Define GC measurement error
Measurement error will be defined as standard deviation of repeated analysis of analytical standards throughout the period of time that the NLA2017 samples were analyzed.

@@

n2o_sd <- 0.007872511

@@

## Simulate measurement error in GC measurement of N2O in equilibrated headspace
```{r simulate_me_Ce, message=F}
C_me <- rep( NA, 1e4 ) # "me" is "measured" gas in headspace. empty vector

# this simulates measurement error on GC.
# assume TRUE value is C_e and error is simulated with with sd of replicate GC measurements
C_e_me <- rnorm( 1e4, C_e, n2o_sd ) # me in measured gas in headspace.
                                   # add random error to measured concentration

for( i in 1:length( C_me ) ){
  C_me[ i ] <- 1e-6 * B * ( V_g * ( C_e_me[ i ] - C_a ) / ( 8.3144598 * ( T_c + 273.15 ) * V_w ) + H * exp( 2700 * ( 1 / ( T_c + 273.15 ) - 1 / 298.15 ) ) * C_e_me[ i ] )
}

@@

## Simulate measurement error for GC measurement of N2O in air
```{r simulate_me_Ca, message=F}
C_me <- rep( NA, 1e4 )

C_a_me <- rnorm( 1e4, C_a, n2o_sd ) #m.e. in measured gas in air

for( i in 1:length( C_me ) ){
  C_me[ i ] <- 1e-6 * B * ( V_g * ( C_e - C_a_me[ i ] ) / ( 8.3144598 * ( T_c + 273.15 ) * V_w ) + H * exp( 2700 * ( 1 / ( T_c + 273.15 ) - 1 / 298.15 ) ) * C_e )
}

# combine into dataframe for ggplot. qplot, which accepts vectors,
# was deprecated in ggplot 3.4.0
Ca_df <- tibble(C_me = C_me, C_a_me = C_a_me)

ggplot(Ca_df, aes(x = C_a_me )) +  
  geom_histogram( binwidth = 0.0001 ) + 
  geom_vline( xintercept = C_a , color = 'red' ) + 
  ggtitle( "Simulated measurement error in air gas concentration") +
  xlab( "concentration (?)")

@@

## Simulate measurement error in $C_e$ and $C_a$
```{r simulate_me_Ca_Ce, message=F}
C_me <- rep( NA, 1e4 )

for( i in 1:length( C_me ) ){
  C_me[ i ] <- 1e-6 * B * ( V_g * ( C_e_me[ i ] - C_a_me[ i ] ) / ( 8.3144598 * ( T_c + 273.15 ) * V_w ) + H * exp( 2700 * ( 1 / ( T_c + 273.15 ) - 1 / 298.15 ) ) * C_e_me[ i ] )
}

@@

## Simulate measurement error in water temperature
```{r simulate_me_T_c, message=F}
C_me <- rep( NA, 1e4 )

@@

for( i in 1:length( C_me ) ){
  C_me[ i ] <- 1e-6 * B * ( V_g * ( C_e - C_a ) / ( 8.3144598 * ( T_c_me[ i ] + 273.15 ) * V_w ) + H * exp( 2700 * ( 1 / ( T_c_me[ i ] + 273.15 ) - 1 / 298.15 ) ) * C_e )
}

# combine into dataframe for ggplot. qplot, which acceptes vectors,
# was deprecated in ggplot 3.4.0
C_df <- tibble(C_me = C_me, T_c_me = T_c_me)

@@

## Simulate measurement error in barometric pressure
```{r simulate_me_B, message=F}
C_me <- rep( NA, 1e4 )

# How to get standard deviation of barometric pressure?
# The manual reports an "accuracy" of +/-3 mm Hg (+/-0.4kPa)
# take standard deviation of simulated measurements with specified uncertainty
B_sd <- sd(runif(n = 1e4, min = B - 0.4, max = B + 0.4)) # calibration certificate reports an uncertainty of +/-0.058C
B_me <- rnorm( 1e4, B, 0.25 ) #m.e. in measured barometric pressure

for( i in 1:length( C_me ) ){
  C_me[ i ] <- 1e-6 * B_me[ i ] * ( V_g * ( C_e - C_a ) / ( 8.3144598 * ( T_c + 273.15 ) * V_w ) + H * exp( 2700 * ( 1 / ( T_c + 273.15 ) - 1 / 298.15 ) ) * C_e )
}

# combine into dataframe for ggplot. qplot, which accepts vectors,
# was deprecated in ggplot 3.4.0
C_df <- tibble(C_me = C_me, B_me = B_me, B_ru = runif(n = 1e4, min = B - 0.4, max = B + 0.4))

@@

ggplot(C_df, aes( x = B_ru )) +  
  geom_histogram( binwidth = 0.01 ) + 
  geom_vline( xintercept = B , color = 'red' ) + 
  ggtitle( "Simulated measurement error in barometric pressure") +
  xlab( "barometric pressure (kPa)")

@@

## Simulate measurement error in water volume
```{r simulate_me_V_w, message=F}
C_me <- rep( NA, 1e4 )

@@

for( i in 1:length( C_me ) ){
  C_me[ i ] <- 1e-6 * B * ( V_g_me[ i ] * ( C_e - C_a ) / ( 8.3144598 * ( T_c + 273.15 ) * V_w_me[ i ] ) + H * exp( 2700 * ( 1 / ( T_c + 273.15 ) - 1 / 298.15 ) ) * C_e )
}

@@

## Simulate measurement error for all observables
```{r simulate_me_all, message=F}
C_me <- rep( NA, 1e4 )

for( i in 1:length( C_me ) ){
  C_me[ i ] <- 1e-6 * B_me[ i ] * ( V_g_me[ i ] * ( C_e_me[ i ] - C_a_me[ i ] ) / ( 8.3144598 * ( T_c_me[ i ] + 273.15 ) * V_w_me[ i ] ) + H * exp( 2700 * ( 1 / ( T_c_me[ i ] + 273.15 ) - 1 / 298.15 ) ) * C_e )
}

@@

## Incorporate measurement uncertainty into source/sink status
The standard deviation of dissolved gas concentration measurements is `r sd(C_me* 1e9)`nM which compares well to 1.25nM reported by Webb et al.  The 95% confidence interval is calculated as +/-2sd, which in this case is `r sd(C_me* 1e9) * 1.96`nM.  Therefore any dissolved gas concentration that is within `r sd(C_me* 1e9) * 1.96`nM of the equilibrium value cannot be determined as a source or a sink.

@@

## Map
ggplot() +
    geom_sf(data = ecoR, color = NA, aes(fill = WSA9_NAME)) +
    geom_sf(data = dg.sf %>% 
              filter(!is.na(n2o.src.snk), sitetype == "PROB", visit.no == 1) %>%
              arrange(desc(n2o.src.snk.with.error)), # grey on top of black
            aes(size = dissolved.n2o.nmol, color = n2o.src.snk.with.error),
            show.legend = "point") +
    geom_sf(data = states, fill = NA, color = "cornsilk3", size = 0.1) +
    # guide argument below removes points from the boxes in the ecoregion legend
    # https://aosmith.rbind.io/2020/07/09/ggplot2-override-aes/
    scale_fill_manual("Ecoregion", values = cols,
                      guide = guide_legend(override.aes = list(shape = NA))) +
    scale_color_manual(values = c("grey", "black", "orange"), name = "source/sink") +
    scale_size(name = expression(N[2]*O~(nM)),
               range = c(0.1, 10), # custom size range
               breaks = c(1, 10, 25, 50, 100)) + # custom breaks
    theme(legend.key.size = unit(0.4, "cm"), # size of boxes in legend
          legend.title = element_text(size = 8))
```
# Set up model for simulating true saturation concentration

@@

$$C_{sat} = 1^{-6}B*C_{air}*(H^{\theta}*exp(2700* \frac{1}{Tc} -  \frac{1}{298.15}))$$
```{r c_sat}
C_sat <- C_a * B * 1e-6 * (H  * exp(2700*(1/(T_c + 273.15) - 1/298.15)))

@@

## Simulate measurement error in GC measurement of N2O in air
```{r simulate_me_sat_C_a, message=F}
C_me_sat <- rep( NA, 1e4 )

C_a_me <- rnorm( 1e4, C_a, n2o_sd ) #m.e. in measured gas in air

for( i in 1:length( C_me_sat ) ){
    C_me_sat[i] <- C_a_me[i] * B * 1e-6 * (H  * exp(2700*(1/(T_c + 273.15) - 1/298.15)))
}

# combine into dataframe for ggplot. qplot, which accepts vectors,
# was deprecated in ggplot 3.4.0
C_sat_df <- tibble(C_me_sat = C_me_sat, C_a_me = C_a_me)

ggplot(C_sat_df, aes(x = C_a_me )) +  
  geom_histogram( binwidth = 0.0001 ) + 
  geom_vline( xintercept = C_a , color = 'red' ) + 
  ggtitle( "Simulated measurement error in air gas concentration") +
  xlab( "concentration (ppm)")

@@

## Simulate measurement error in water temperature
```{r simulate_me_sat_T_c, message=F}
C_me_sat <- rep( NA, 1e4 )

for( i in 1:length( C_me_sat ) ){
    C_me_sat[i] <- C_a * B * 1e-6 * (H  * exp(2700*(1/(T_c_me[ i ] + 273.15) - 1/298.15)))
}

# combine into dataframe for ggplot. qplot, which acceptes vectors,
# was deprecated in ggplot 3.4.0
C_sat_df <- tibble(C_me_sat = C_me_sat, T_c_me = T_c_me)

@@

## Simulate measurement error in barometric pressure
```{r simulate_me_sat_B, message=F}
C_me_sat <- rep( NA, 1e4 )

for( i in 1:length( C_me_sat ) ){
    C_me_sat[i] <- C_a * B_me[ i ] * 1e-6 * (H  * exp(2700*(1/(T_c + 273.15) - 1/298.15)))
}

@@

## Simulate measurement error for all observables
```{r simulate_me_sat_all, message=F}
C_me_sat <- rep( NA, 1e4 )

for( i in 1:length( C_me_sat ) ){
    C_me_sat[i] <- C_a_me[ i ] * B_me[ i ] * 1e-6 * (H  * exp(2700*(1/(T_c_me[ i ] + 273.15) - 1/298.15)))
}

@@

``` {r}

C_ratio <- (1e-6 * B * ( V_g * ( C_e - C_a ) / ( 8.3144598 * ( T_c + 273.15 ) * V_w ) + H * exp( 2700 * ( 1 / ( T_c + 273.15 ) - 1 / 298.15 ) ) * C_e )) /
  (C_a * B * 1e-6 * (H  * exp(2700*(1/(T_c + 273.15) - 1/298.15))))

```
## Simulate measurement error for all observables
```{r simulate_me_ratio_all, message=F}
C_me_ratio <- rep( NA, 1e4 )

for( i in 1:length( C_me_ratio ) ){
C_me_ratio[ i ] <- (1e-6 * B_me[i] * ( V_g_me[i] * ( C_e_me[i] - C_a_me[i] ) / ( 8.3144598 * ( T_c_me[i] + 273.15 ) * V_w_me[i] ) + H * exp( 2700 * ( 1 / ( T_c_me[i] + 273.15 ) - 1 / 298.15 ) ) * C_e_me[i] )) /
  (C_a_me[i] * B_me[i] * 1e-6 * (H  * exp(2700*(1/(T_c_me[i] + 273.15) - 1/298.15)))) 
}

@@

# standard deviation of a saturation ratio value s:
sd(C_me_ratio) # 0.0.05479

# the 95% CI is therefore the computed value +/- 2 standard deviations
sd(C_me_ratio) * 1.96 # +/- 0.1073

# so any ratio within 0.1073 of 1.0 cannot be discerned as a source or sink?
```

>>>>

```{=html}
<style>
.vscroll-plot {
    width: 800px;
    height: 600px;
    overflow-y: scroll;
    overflow-x: hidden;
}
</style>
```
# Background

In this document we explore a range of models for fitting dissolved $\text{N}_2\text{O}$ data observed as part of the 2017 National Lakes Assessment (NLA).

# Setup R

```{r setup_1, include=FALSE, echo=FALSE}
library(knitr)
# opts_knit$set(root.dir = "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/scripts")
# 
# #set directories and R package library
# setwd("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/scripts")
# .libPaths("C:/Users/rmartin/Desktop/R/library")
```

@@

# Set up model for simulating true dissolved N2O value

@@

B <- 99 # barometric pressure (kPa)
V_g <- 35 # volume of reference gas in headspace
V_w <- 105 # volume of water below headspace
T_c <- 23 # temperature ( C )
C_e <- 0.307 # gas mixture ratio in headspace
C_a_n2o <- 0.310 # gas mixture ratio in reference
H_n2o <- 0.00024 #mol m-3 Pa, range: 0.00018 - 0.00025; Henry's law constant
H_ar <- 0.000014 #mol m-3 Pa, range: 0.000013 - 0.000016; Henry's law constant

C <- 1e-6 * B * ( V_g * ( C_e - C_a_n2o ) / ( 8.3144598 * ( T_c + 273.15 ) * V_w ) + H_n2o * exp( 2700 * ( 1 / ( T_c + 273.15 ) - 1 / 298.15 ) ) * C_e )

@@

## Define GC measurement error

Measurement error will be defined as standard deviation of repeated analysis of analytical standards throughout the period of time that the NLA2017 samples were analyzed.

@@

n2o_sd_epa <- 0.007872511 # from AWBERC GC

@@

## Simulate measurement error in GC measurement of N2O in equilibrated headspace

```{r simulate_me_Ce, message=F}
C_me <- rep( NA, 1e4 ) # "me" is "measured" gas in headspace. empty vector

# this simulates measurement error on GC.
# assume TRUE value is C_e and error is simulated with with sd of replicate GC measurements
C_e_me <- rnorm( 1e4, C_e, n2o_sd_epa ) # me in measured gas in headspace.
                                   # add random error to measured concentration

for( i in 1:length( C_me ) ){
  C_me[ i ] <- 1e-6 * B * ( V_g * ( C_e_me[ i ] - C_a_n2o ) / ( 8.3144598 * ( T_c + 273.15 ) * V_w ) + H_n2o * exp( 2700 * ( 1 / ( T_c + 273.15 ) - 1 / 298.15 ) ) * C_e_me[ i ] )
}

@@

## Simulate measurement error for GC measurement of N2O in air

```{r simulate_me_Ca, message=F}
C_me <- rep( NA, 1e4 )

C_a_n2o_me <- rnorm( 1e4, C_a_n2o, n2o_sd_epa ) #m.e. in measured gas in air

for( i in 1:length( C_me ) ){
  C_me[ i ] <- 1e-6 * B * ( V_g * ( C_e - C_a_n2o_me[ i ] ) / ( 8.3144598 * ( T_c + 273.15 ) * V_w ) + H_n2o * exp( 2700 * ( 1 / ( T_c + 273.15 ) - 1 / 298.15 ) ) * C_e )
}

# combine into dataframe for ggplot. qplot, which accepts vectors,
# was deprecated in ggplot 3.4.0
Ca_df <- tibble(C_me = C_me, C_a_n2o_me = C_a_n2o_me)

ggplot(Ca_df, aes(x = C_a_n2o_me )) +  
  geom_histogram( binwidth = 0.0001 ) + 
  geom_vline( xintercept = C_a_n2o , color = 'red' ) + 
  ggtitle( "Simulated measurement error in air gas concentration") +
  xlab( "concentration (?)")

@@

## Simulate measurement error in $C_e$ and $C_a_n2o$

```{r simulate_me_Ca_Ce, message=F}
C_me <- rep( NA, 1e4 )

for( i in 1:length( C_me ) ){
  C_me[ i ] <- 1e-6 * B * ( V_g * ( C_e_me[ i ] - C_a_n2o_me[ i ] ) / ( 8.3144598 * ( T_c + 273.15 ) * V_w ) + H_n2o * exp( 2700 * ( 1 / ( T_c + 273.15 ) - 1 / 298.15 ) ) * C_e_me[ i ] )
}

@@

## Simulate measurement error in water temperature

```{r simulate_me_T_c, message=F}
C_me <- rep( NA, 1e4 )

@@

for( i in 1:length( C_me ) ){
  C_me[ i ] <- 1e-6 * B * ( V_g * ( C_e - C_a_n2o ) / ( 8.3144598 * ( T_c_me[ i ] + 273.15 ) * V_w ) + H_n2o * exp( 2700 * ( 1 / ( T_c_me[ i ] + 273.15 ) - 1 / 298.15 ) ) * C_e )
}

# combine into dataframe for ggplot. qplot, which accepts vectors,
# was deprecated in ggplot 3.4.0
C_df <- tibble(C_me = C_me, T_c_me = T_c_me)

@@

## Simulate measurement error in barometric pressure

```{r simulate_me_B, message=F}
C_me <- rep( NA, 1e4 )

# How to get standard deviation of barometric pressure?
# The manual reports an "accuracy" of +/-3 mm Hg (+/-0.4kPa), but EPA calibration
# checks performed in the AWBERC metrology lab demonstrate much greater precision.
# below we calculat sd of replicate measurements made with 15 different barometers.
# we then take the mean of the 15 sd values.

#2023 barometer verification in AWBERC metrology lab
B_sd <- mean(
  sd(c(999,998.6)*0.1), # YSI ProDSS
  sd(c(1000.848, 1000.848)*0.1), # EXO 21G104022
  sd(c(1000.982, 1000.982)*0.1), # EXO 18H111572
  sd(c(1000.982, 1000.982)*0.1), # EXO 22F105975
  sd(c(1001.382, 1001.382)*0.1), # EXO 21K100162
  sd(c(1000.98, 1000.85)*0.1), # EXO 21F103579
  sd(c(1000.31, 1000.44)*0.1), # MDS 16D100327
  sd(c(1001.25, 1001.25)*0.1), # MDS 12F102618
  sd(c(1000, 1000)*0.1), # MDS 16D100328
  sd(c(1002, 1002)*0.1), # MDS 02K599 AF
  sd(c(1003,1003)*0.1), # MDS 12J0793 AF
  sd(c(1000.85,1000.85)*0.1), # MDS 16E100831
  sd(c(1000.32, 1000.32)*0.1), # MDS 12F102619
  sd(c(1002, 1002)*0.1)) # MDS 15F104725

B_me <- rnorm( 1e4, B, B_sd ) #m.e. in measured barometric pressure

for( i in 1:length( C_me ) ){
  C_me[ i ] <- 1e-6 * B_me[ i ] * ( V_g * ( C_e - C_a_n2o ) / ( 8.3144598 * ( T_c + 273.15 ) * V_w ) + H_n2o * exp( 2700 * ( 1 / ( T_c + 273.15 ) - 1 / 298.15 ) ) * C_e )
}

# combine into dataframe for ggplot. qplot, which accepts vectors,
# was deprecated in ggplot 3.4.0
C_df <- tibble(C_me = C_me, B_me = B_me)

@@

@@

## Simulate measurement error in water volume

```{r simulate_me_V_w, message=F}
C_me <- rep( NA, 1e4 )

@@

for( i in 1:length( C_me ) ){
  C_me[ i ] <- 1e-6 * B * ( V_g_me[ i ] * ( C_e - C_a_n2o ) / ( 8.3144598 * ( T_c + 273.15 ) * V_w_me[ i ] ) + H_n2o * exp( 2700 * ( 1 / ( T_c + 273.15 ) - 1 / 298.15 ) ) * C_e )
}

@@

## Simulate measurement error for all observables

```{r simulate_me_all, message=F}
C_me <- rep( NA, 1e4 )

for( i in 1:length( C_me ) ){
  C_me[ i ] <- 1e-6 * B_me[ i ] * ( V_g_me[ i ] * ( C_e_me[ i ] - C_a_n2o_me[ i ] ) / ( 8.3144598 * ( T_c_me[ i ] + 273.15 ) * V_w_me[ i ] ) + H_n2o * exp( 2700 * ( 1 / ( T_c_me[ i ] + 273.15 ) - 1 / 298.15 ) ) * C_e )
}

@@

## Incorporate measurement uncertainty into source/sink status

The standard deviation of dissolved gas concentration measurements is `r sd(C_me* 1e9)`nM which compares well to 1.25nM reported by Webb et al. The 95% confidence interval is calculated as +/-2sd, which in this case is `r sd(C_me* 1e9) * 1.96`nM. Therefore any dissolved gas concentration that is within `r sd(C_me* 1e9) * 1.96`nM of the equilibrium value cannot be determined as a source or a sink.

@@

## Map
ggplot() +
    geom_sf(data = ecoR, color = NA, aes(fill = WSA9_NAME)) +
    geom_sf(data = dg.sf %>% 
              filter(!is.na(n2o.src.snk), sitetype == "PROB", visit.no == 1) %>%
              arrange(desc(n2o.src.snk.with.error)), # grey on top of black
            aes(size = dissolved.n2o.nmol, color = n2o.src.snk.with.error),
            show.legend = "point") +
    geom_sf(data = states, fill = NA, color = "cornsilk3", size = 0.1) +
    # guide argument below removes points from the boxes in the ecoregion legend
    # https://aosmith.rbind.io/2020/07/09/ggplot2-override-aes/
    scale_fill_manual("Ecoregion", values = cols,
                      guide = guide_legend(override.aes = list(shape = NA))) +
    scale_color_manual(values = c("grey", "black", "orange"), name = "source/sink") +
    scale_size(name = expression(N[2]*O~(nM)),
               range = c(0.1, 10), # custom size range
               breaks = c(1, 10, 25, 50, 100)) + # custom breaks
    theme(legend.key.size = unit(0.4, "cm"), # size of boxes in legend
          legend.title = element_text(size = 8))
```

# Set up model for simulating true saturation concentration

@@

$$C_{sat} = 1^{-6}B*C_{air}*(H^{\theta}*exp(2700* \frac{1}{Tc} -  \frac{1}{298.15}))$$

```{r c_sat}
C_sat <- C_a_n2o * B * 1e-6 * (H_n2o  * exp(2700*(1/(T_c + 273.15) - 1/298.15)))

@@

## Simulate measurement error in GC measurement of N2O in air

```{r simulate_me_sat_C_a_n2o, message=F}
C_me_sat <- rep( NA, 1e4 )

C_a_n2o_me <- rnorm( 1e4, C_a_n2o, n2o_sd_epa ) #m.e. in measured gas in air

for( i in 1:length( C_me_sat ) ){
    C_me_sat[i] <- C_a_n2o_me[i] * B * 1e-6 * (H_n2o  * exp(2700*(1/(T_c + 273.15) - 1/298.15)))
}

# combine into dataframe for ggplot. qplot, which accepts vectors,
# was deprecated in ggplot 3.4.0
C_sat_df <- tibble(C_me_sat = C_me_sat, C_a_n2o_me = C_a_n2o_me)

ggplot(C_sat_df, aes(x = C_a_n2o_me )) +  
  geom_histogram( binwidth = 0.0001 ) + 
  geom_vline( xintercept = C_a_n2o , color = 'red' ) + 
  ggtitle( "Simulated measurement error in air gas concentration") +
  xlab( "concentration (ppm)")

@@

## Simulate measurement error in water temperature

```{r simulate_me_sat_T_c, message=F}
C_me_sat <- rep( NA, 1e4 )

for( i in 1:length( C_me_sat ) ){
    C_me_sat[i] <- C_a_n2o * B * 1e-6 * (H_n2o  * exp(2700*(1/(T_c_me[ i ] + 273.15) - 1/298.15)))
}

# combine into dataframe for ggplot. qplot, which accepts vectors,
# was deprecated in ggplot 3.4.0
C_sat_df <- tibble(C_me_sat = C_me_sat, T_c_me = T_c_me)

@@

## Simulate measurement error in barometric pressure

```{r simulate_me_sat_B, message=F}
C_me_sat <- rep( NA, 1e4 )

for( i in 1:length( C_me_sat ) ){
    C_me_sat[i] <- C_a_n2o * B_me[ i ] * 1e-6 * (H_n2o  * exp(2700*(1/(T_c + 273.15) - 1/298.15)))
}

@@

## Simulate measurement error for all observables

```{r simulate_me_sat_all, message=F}
C_me_sat <- rep( NA, 1e4 )

for( i in 1:length( C_me_sat ) ){
    C_me_sat[i] <- C_a_n2o_me[ i ] * B_me[ i ] * 1e-6 * (H_n2o  * exp(2700*(1/(T_c_me[ i ] + 273.15) - 1/298.15)))
}

@@

```{r}

C_ratio <- (1e-6 * B * ( V_g * ( C_e - C_a_n2o ) / ( 8.3144598 * ( T_c + 273.15 ) * V_w ) + H_n2o * exp( 2700 * ( 1 / ( T_c + 273.15 ) - 1 / 298.15 ) ) * C_e )) /
  (C_a_n2o * B * 1e-6 * (H_n2o  * exp(2700*(1/(T_c + 273.15) - 1/298.15))))

## Simulate measurement error for all observables when air is used as headspace gas

```{r simulate_me_ratio_all, message=F}
C_me_ratio <- rep( NA, 1e4 )

for( i in 1:length( C_me_ratio ) ){
C_me_ratio[ i ] <- (1e-6 * B_me[i] * ( V_g_me[i] * ( C_e_me[i] - C_a_n2o_me[i] ) / ( 8.3144598 * ( T_c_me[i] + 273.15 ) * V_w_me[i] ) + H_n2o * exp( 2700 * ( 1 / ( T_c_me[i] + 273.15 ) - 1 / 298.15 ) ) * C_e_me[i] )) /
  (C_a_n2o_me[i] * B_me[i] * 1e-6 * (H_n2o  * exp(2700*(1/(T_c_me[i] + 273.15) - 1/298.15)))) 
}

@@

# standard deviation of a saturation ratio value s:
sd(C_me_ratio) # 0.026

# the 95% CI is therefore the computed value +/- 2 standard deviations
sd(C_me_ratio) * 1.96 # +/- 0.05258214

# so any ratio within 0.05 of 1.0 cannot be discerned as a source or sink?
```

## Map of saturation ratio values with uncertainty when air is used as headspace gas

```{r map_sat_ratio, message=F}
# two standard deviations = 0.3019449
# 95% CI is therefore x+/-0.3019449

dg <- dg %>%
  mutate(n2o.src.snk.with.all.error = case_when(
    abs(n2o.sat.ratio - 1) < sd(C_me_ratio) ~ "undetermined",
    TRUE ~ n2o.src.snk))

dg %>% 
  filter(sample.source == "DG", visit.no == 1, sitetype == "PROB") %>%
  {table(.$n2o.src.snk.with.all.error)} %>%
  kable(digits = 2, caption = "Number of lakes classified as source/sink")

dg %>% 
  filter(sample.source == "DG", visit.no == 1, sitetype == "PROB") %>%
  {prop.table(table(.$n2o.src.snk.with.all.error))} %>%
  kable(digits = 2, caption = "Proportion of lakes classified as source/sink")

# sf for plotting
dg.sf <- st_as_sf(dg, coords = c("map.lon.dd", "map.lat.dd"), 
                  crs = 4269) %>% # standard for lat/lon
  st_transform(5070) # project to CONUS ALBERS for plotting

## Map
ggplot() +
    geom_sf(data = ecoR, color = NA, aes(fill = WSA9_NAME)) +
    geom_sf(data = dg.sf %>% 
              filter(!is.na(n2o.src.snk), sitetype == "PROB", visit.no == 1) %>%
              arrange(desc(n2o.src.snk.with.all.error)), # grey on top of black
            aes(size = dissolved.n2o.nmol, color = n2o.src.snk.with.all.error),
            show.legend = "point") +
    geom_sf(data = states, fill = NA, color = "cornsilk3", size = 0.1) +
    # guide argument below removes points from the boxes in the ecoregion legend
    # https://aosmith.rbind.io/2020/07/09/ggplot2-override-aes/
    scale_fill_manual("Ecoregion", values = cols,
                      guide = guide_legend(override.aes = list(shape = NA))) +
    scale_color_manual(values = c("grey", "black", "orange"), name = "source/sink") +
    scale_size(name = expression(N[2]*O~(nM)),
               range = c(0.1, 10), # custom size range
               breaks = c(1, 10, 25, 50, 100)) + # custom breaks
    theme(legend.key.size = unit(0.4, "cm"), # size of boxes in legend
          legend.title = element_text(size = 8))
```

## Simulate measurement error for all observables when a pure gas (e.g. helium, nitrogen) is used as headspace gas

```{r simulate_me_ratio_pure_gas_all, message=F}
# set C_e to a lower value to generate a saturation ratio close to that used above
# C_e will be lower due to dilution with pure headspace gas (e.g. helium, nitrogen)
C_e <- 0.2

# this simulates measurement error on GC.
# assume TRUE value is C_e and error is simulated with with sd of replicate GC measurements
C_e_me <- rnorm( 1e4, C_e, n2o_sd_epa ) # me in measured gas in headspace.
                                   # add random error to measured concentration

# under this scenario we don't need to subtract C_a_n2o (N2O concentration in air) from C_e (N2O concentration in equilibrated headspace)
C_ratio <- (1e-6 * B * ( V_g * ( C_e - 0 ) / ( 8.3144598 * ( T_c + 273.15 ) * V_w ) + H_n2o * exp( 2700 * ( 1 / ( T_c + 273.15 ) - 1 / 298.15 ) ) * C_e )) /
  (C_a_n2o * B * 1e-6 * (H_n2o  * exp(2700*(1/(T_c + 273.15) - 1/298.15))))

C_me_ratio <- rep( NA, 1e4 )

for( i in 1:length( C_me_ratio ) ){
C_me_ratio[ i ] <- (1e-6 * B_me[i] * ( V_g_me[i] * ( C_e_me[i] - 0 ) / ( 8.3144598 * ( T_c_me[i] + 273.15 ) * V_w_me[i] ) + H_n2o * exp( 2700 * ( 1 / ( T_c_me[i] + 273.15 ) - 1 / 298.15 ) ) * C_e_me[i] )) /
  (C_a_n2o_me[i] * B_me[i] * 1e-6 * (H_n2o  * exp(2700*(1/(T_c_me[i] + 273.15) - 1/298.15)))) 
}

# standard deviation of a saturation ratio value s:
sd(C_me_ratio) # 0.048

# the 95% CI is therefore the computed value +/- 2 standard deviations
sd(C_me_ratio) * 1.96 # +/- 0.094

# so any ratio within 0.094 of 1.0 cannot be discerned as a source or sink
```

## Simulate measurement error for all observables when a pure gas (e.g. helium, nitrogen) is used as headspace gas combined with high precision GC

The GC used in the current study is optimized to quantify several analytes across a broad range of concentrations but greater precision can be achieved by optimizing the instrument for a single analyte across a narrower range of concentrations.  For example,  laboratories participating in an international analysis of N2O standards used for atmsopheric monitoring reported a mean standard deviation of `r mean(c(.00033, .00034, .00031, 0.00030, 0.0016, 0.0015))`, which is 10-fold lower than the standard deviation of N2O measurements in this study (`r n2o_sd_epa`).  Utilizing more precise N2O measurements for the equilibrated headspace and/or air will reduce uncertainty in the estimated dissolved N2O concentration, dissolved N2O equilibrium value, and N2O saturation ratio.

```{r simulate_me_ratio_gc_pure_gas_all, message=F}
# set C_e to a lower value to generate a saturation ratio close to that used above
# C_e will be lower due to dilution with pure headspace gas (e.g. helium, nitrogen)
C_e <- 0.2

# two standard deviations reported from each of three labs participating in
# a laboratory intercomparison study
n2o_sd_lit <- mean(c(.00033, .00034, .00031, 0.00030, 0.0016, 0.0015)) # from literature

# simulate improved measurement of N2O in equilibrated headspace gas 
# assume TRUE value is C_e and error is simulated with with sd from literature
C_e_me <- rnorm( 1e4, C_e, n2o_sd_lit )

# under this scenario we don't need to subtract C_a_n2o (N2O concentration in air) from C_e (N2O concentration in equilibrated headspace)
C_ratio <- (1e-6 * B * ( V_g * ( C_e - 0 ) / ( 8.3144598 * ( T_c + 273.15 ) * V_w ) + H_n2o * exp( 2700 * ( 1 / ( T_c + 273.15 ) - 1 / 298.15 ) ) * C_e )) /
  (C_a_n2o * B * 1e-6 * (H_n2o  * exp(2700*(1/(T_c + 273.15) - 1/298.15))))

C_me_ratio <- rep( NA, 1e4 )

for( i in 1:length( C_me_ratio ) ){
C_me_ratio[ i ] <- (1e-6 * B_me[i] * ( V_g_me[i] * ( C_e_me[i] - 0 ) / ( 8.3144598 * ( T_c_me[i] + 273.15 ) * V_w_me[i] ) + H_n2o * exp( 2700 * ( 1 / ( T_c_me[i] + 273.15 ) - 1 / 298.15 ) ) * C_e_me[i] )) /
  (C_a_n2o_me[i] * B_me[i] * 1e-6 * (H_n2o  * exp(2700*(1/(T_c_me[i] + 273.15) - 1/298.15)))) 
}

# standard deviation of a saturation ratio value s:
sd(C_me_ratio) # 0.029

# the 95% CI is therefore the computed value +/- 2 standard deviations
sd(C_me_ratio) * 1.96 # +/- 0.0569

# so any ratio within 0.0569 of 1.0 cannot be discerned as a source or sink
```

# Set up model for MIMS

In Membrane Inlet Mass Spectrometry (MIMS), dissolved gases (typically N2, Ar, O2, N2O) are extracted from a water sample as it passes through a semi-permeable membrane. The abundance of the extracted gases are then measured using a quadrapole mass spectrometer. One advantage of this approach relative to the traditional headspace equilibrium method is that dissolved gases are extracted without contamination and/or dilution from a headspace gas (e.g. air, nitrogen, helium). Dilution by a pure gas isn't much of a concern because N2O detection limits are very low via GC, but correction for the presence of N2O in the headspace gas reduces measurement precision.

Dissolved N2O concentration in a MIMS sample can be calculated from the N2O signal strength and a calibration factor. The calibration factor is determined by analyzing a water sample with known N2O concentration. This is most frequently accomplished by equilibrating laboratory water with air at constant temperature, barometric pressure and N2O partial pressure. The dissolved N2O concentration of the equilibrated water is calculated from solubility laws (see [Set up model for simulating true saturation concentration]) and the calibration factor is:

$$N_2O \text{ calibration factor} =\frac{[N_2O_{\text{dissolved standard}}]}{N_2O_{\text{standard}} \text{ signal strength}}$$

The N2O concentration of an unknown sample is then calculated as:

$$C = [N_2O_{sample}] = N_2O_{\text{sample}} \text{ signal strength} * N_2O \text{ calibration factor}$$
